The evolution of the transcriptome of mammalian placentas

### 2016-11-15 manuscript/manuscript.Rnw
discuss annexin in the context of cell fusion

Another annexin which is highly expressed in the core placenta is
\textit{ANXA5}, a gene which was originally identified as

@@

>>>>

Another annexin which is highly expressed in all term placentas is
\textit{ANXA5}, which was originally identified as a an anti-coagulant
protein, but is also involved in syncitium formation, membrane repair,
and reduction of spontaneous
abortion~\cite{Ueki.ea2012:Lossofmaternalannexin}. Other notable
syncytiotrophoblast fusion proteins, Syncytin-1 (\textit{ERVW-1}) and -2
(\textit{ERVFRD-1}) were not present in the core placenta, as they (or
their homologs) are not highly expressed in \dog, \bonobo, or \sheep,
nor was \textit{MFSD2A}, the Syncytin-2 receptor (\ac{FPKM} $\ge 10$
only in \human, \mouse, and \dog).

@@

\subsubsection{Formation of the syncytiotrophoblast}

While Syncitin-1 and -2 are implicated in the formation of the
syncytiotrophoblast in
\human~\cite{Rawn.ea2008:evolutionregulationandfunction} and homologs
of \textit{ERVFRD-1} and \textit{ERVW-1} are widely expressed in most
of the placentas analyzed here, they are not expressed in \bonobo,
\sheep, and \dog, even though these mammals form
syncytiotrophoblasts~\cite{Mess.ea2006:Evolutionarytransformationsoffetal}.
Because the formation of syncytiotrophoblast is the stem state of
Eutheria, Afrotheria, Euarchontoglires, and
Laurasiatheria~\cite{Mess.ea2006:Evolutionarytransformationsoffetal},
the gene(s) responsible for cytotrophoblast fusion to form
syncytiotrophoblast is likely to be highly expressed in the core
placenta. \textit{ANXA5} is known to be expressed at multiple stages
of placentation, and is capable of repairing membrane defects and
promoting the fusion of cells~\DLA{Cite more papers here which define
  this.} in the presence of Ca$^{2+}$. While there are likely other
genes which are responsible for the positioning of cytotrophoblast
cells in close proximity to start the fusion process, \textit{ANXA5}
is likely to be essential for the completion of fusion.

\DLA{Possibility of membrane repair due to maternal immune response as
  well during this process, as well as repair of membrane due to high
  secretion and adsoroption of nutrients/waste in the placenta.}

% Cows have BERV-K1 and -K2, but it's not annotated in the ENSEMBL
% build of the cow genome that we have
